---
title: "case_study"
author: "Kimia Faroughi"
date: "2025-10-07"
output: github_document
---

```{r setup, echo = FALSE, message = FALSE}
library(tidyverse)
library(patchwork)

#global ggplot settings
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

Load key packages.

```{r}
library(p8105.datasets)
library(leaflet)

data(nyc_airbnb)
```

## Brainstorm questions

 * Pick a neighborhood / borough, investigate price
    * If we had dates, does that affect price?
 * Does the type of unit affect rating / price?
 * What about ZIP codes?
 * What rental has the highest reviews?
 * Are there any connections between price and other variables?
 * What about prices over time?
    * Could we look at prices before and after the pandemic?
 * How else did the pandemic affect airbnb in nyc?
 * Does the name of the rental affect the price / rating?
 * What does the number of days available have to do with anything else?
 
Let's look at prices first

```{r}
nyc_airbnb |> 
  drop_na(price) |> 
  group_by(neighbourhood_group) |> 
  summarize(
    mean_price = mean(price),
    min_price = min(price),
    max_price = max(price)
  )
```

What about room type in a given neighborhood?

```{r}
nyc_airbnb |> 
  drop_na(price) |> 
  group_by(room_type, neighbourhood_group) |> 
  summarize(
    mean_price = mean(price)
  ) |> 
  pivot_wider(
    names_from = room_type,
    values_from = mean_price
  )
```

```{r}
nyc_airbnb |> 
  count(neighbourhood_group, room_type)
```

Make some plots

```{r}
nyc_airbnb |> 
  ggplot(aes(x = price)) +
  geom_histogram() +
  facet_grid(room_type ~ neighbourhood_group) + 
  scale_x_continuous(limits = c(0, 1000)) #zooms in on x-axis scale
```

```{r}
nyc_airbnb |> 
  filter(neighbourhood_group == "Manhattan",
         price < 500,
         room_type == "Entire home/apt") |> 
  mutate(neighbourhood = fct_reorder(neighbourhood, price)) |> #goes from smallest to largest
  ggplot(aes(x = neighbourhood, y = price)) + 
  geom_violin() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) #makes axis text more readable
```

(remember to look at extreme values)

```{r}
nyc_airbnb |> 
  filter(price > 9000)
```

```{r}
nyc_airbnb |> 
  filter(calculated_host_listings_count > 30)
```

Keep looking at location 

```{r}
nyc_airbnb |> 
  filter(neighbourhood_group == "Manhattan",
         price < 500,
         room_type == "Entire home/apt") |> 
  ggplot(aes(x = lat, y = long, color = review_scores_location)) +
  geom_point()
```

Can we make a better map?

```{r}
nyc_airbnb |> 
  filter(neighbourhood_group == "Manhattan",
         price < 500,
         room_type == "Entire home/apt") |> 
  leaflet() |> 
  addTiles() |> 
  addMarkers(~lat, ~long)
```

